The Essence of Summarization\\ in Pushdown Flow Analyses\\ \textbf{Supplemental Material}

### 2013-12-13 content.tex
Shrinking.

Programs in higher-order languages heavily use function calls and method dispatch as part of their control flow.
%
Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function or method calls and subsequent returns.
%
These works, called CFA2 and PDCFA respectively, use pushdown automata as their approximation's target model of computation.
%
They are hence called ``pushdown analyses.''
%
\footnote{We will refer to the classic finite model analyses as ``regular analyses'' after regular languages.}
%
CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics --- especially if they feature non-local control transfer that breaks the pushdown model.
%%

@@

%%
The $\Addr$ space is what controls the precision of the model.
%
For a concrete semantics, we require the allocation meta-function $\alloc :
\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
%
If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of continuation frames is finite.
%
A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet'' --- thus the pushdown system interpretation is apt.
%
The $\alloc$ meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
%
All recursion can be expressed by indirecting through addresses into a store.
%
We borrow this technique on top of our own.

@@

%%
The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
%
A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
%
In this case, the context includes the function, the environment, and the store.
%
All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
%%

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores --- this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified thusly:

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

\begin{figure}
      \begin{tabular}{rcl}
        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
        & $\Rightarrow$ &
        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$ \\
        & & \multicolumn{1}{c}{$\Downarrow$} \\
        \begin{minipage}{0.5\linewidth}
          \begin{center}
            {\small
              \begin{tabular}{|l|c|c|c|}
                \hline
                \# & $\mktab$ & $\mmemo$ \\
                \hline
                0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
                1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
                \hline
              \end{tabular}} \\[8pt]
            {\small
              \begin{tabular}{|l|c|}
                \hline
                \# & $\mctx$ \\
                \hline
                0  & $(({\tt x}, \iota), \mstore_1)$ \\
                1  & $(({\tt x}, \iota), \mstore_3)$ \\
                \hline
              \end{tabular}}
          \end{center}

      \end{minipage} & &
        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
      \end{tabular}
    \begin{center} {\small
        \begin{tabular}{|l|c|c|}
          \hline
          \# & $\mstore$ & $\mkont$ \\
          \hline
          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
          1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
          3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
          4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
          \hline
        \end{tabular}} \\[10pt]
 Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
    \end{center}
\caption{Example analysis evaluation}
\label{fig:ex-eval}
\end{figure}
% %% FIXME stores in memo-table
% \begin{enumerate}
% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
% \end{enumerate}

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$, that is different for our continuations.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{center}
  \,
  \begin{minipage}{0.32\linewidth}
    $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
  \end{minipage}
  \begin{minipage}{0.36\linewidth}
    $\touchesf(\kar{\mexpr,\menv}) = \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})$
  \end{minipage}
  \begin{minipage}{0.3\linewidth}
    $\touchesf(\kfn{\mval}) = \touches(\mval)$
  \end{minipage}
\end{center}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

@@

A tabular semantics for the CES$\xi$K machine follows the previous section almost entirely, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
\begin{align*}
  \tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
  \tpl{\mpoint,
    \mstore',
    \mframe',
    \krt{\mpoint, \mstore'},
    \mktab',
    \mmemo} \\
  \text{where }
  & \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
  & (\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval) \\
  & \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{(\mkont,\mframe)}
\\[2pt]
 \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
 \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
  \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
\end{align*}

%%
CFA2 also includes what they called ``transitive summaries'' to deal with tail calls.
%
We create an $\rt$ continuation at every function call.
%
This view contends with tail calls, but we can identify tail calls easily --- any call with an $\rt$ as its continuation is a tail call.
%
Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important since there is no unbounded recursion.
%
The repeated popping of $\rt$ inserted by what otherwise were tail-calls is synonymous with CFA2's transitive summaries.

@@

Second, the $\domarkset$ metafunction builds a map of keys to lists, and builds a private store for the list constructions.
%
For this, we extend $\alloc$'s specification to allow more arguments --- the private store, and a set of keys that need cons cells allocated for them --- and produce a map from keys to pairs of addresses.
%
%% FIXME: need to crawl through qualified continuations for all extensions to build a list for each found key.
%% Has another implicit equality check
\newcommand{\markaux}{\mathit{marks}}
\begin{align*}
  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &= \markaux(\mkont,\emptyset) \\
  \markaux(\kmt^\bot, G) &= (\bot, \bot) \\
  \markaux(\krt{\mctx}^\mmarks, G) &= (\bot, \bot) \text{ if } \mctx \in G \\ %% FIXME: unsound
  \markaux(\krt{\mctx}^\mmarks, G) &= \bigsqcup\limits_{\mkont\in \mktab(\mctx)}{\markaux(\mkont,G \cup \set{\mctx})} \text{ otherwise} \\
  \markaux(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, G) &= (\mmarks''[\mval_k \mapsto \set{\cons{a}{d}} : \mval_k \in \dom(\mmarks), (a, d) = A(\mval_k)], \\
    &\phantom{= (}
      \mstore\sqcup[a \mapsto \mmarks(\mval_k) : \mval_k \in \dom(\mmarks), (a,\_) = A(\mval_k)] \\
    &\phantom{= (\mstore}\sqcup[d \mapsto \mmarks''(\mval_k) : \mval_k \in \dom(\mmarks), (\_,d) = A(\mval_k)]
   \\ \text{where } (\mmarks'', \mstore) &= \markaux(\mkont, G)
   \\ A &= \alloc(\mstate, \mstore, \dom(\mmarks))
\end{align*}
}

@@

At return boundaries, the memo table must add the result to all the represented contexts:
\begin{align*}
  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \mstore', \mmktab'), \mathit{out}) &=
  [((\mexpr,\menv), \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
    \text{where } C &= \setbuild{((\mexpr,\menv), \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
\end{align*}

@@

\paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
We lose precision if we use a CPS transform to compile away {\tt shift} and {\tt reset} forms, because variables are treated less precisely than continuations.
%
%In pushdown analysis, we track function return points with entire calling contexts rather than the $k$-CFA way of allocating the continuation in the heap at some lower-precision address, like a variable.
%
%In our summarizing analysis of delimited composable control, we get added precision for {\tt reset}, {\tt shift}, and continuation invocation points.
%
Consider the following program and its CPS transform for comparison:
% \begin{lstlisting}[mathescape]
% (let* ([id ($\lambda$ (x) x)]
%        [f ($\lambda$ (y) (shift k (k (k y))))]
%        [g ($\lambda$ (z) (reset (id (f z))))])
%   (<= (g 0) (g 1)))
% \end{lstlisting}
\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}

@@

\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\le$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}

PDCFA with a monovariant allocation strategy will predict the CPS'd version returns true or false.
%
Our specialized analysis for delimited control, will predict the non-CPS'd version returns true.}

@@

%%
Pushdown models have existed in the first-order static analysis literature~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
%
The important difference when we move higher-order is that the model construction to feed these methods is an additional problem --- the one we solve here.
%%

>>>>

Programs in higher-order languages heavily use function calls and method dispatch as part of their control flow.
%
Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function or method calls and subsequent returns.
%
These works, called CFA2 and PDCFA respectively, use pushdown automata as their approximation's target model of computation.
%
They are hence called ``pushdown analyses.''%
%
\footnote{We refer to finite model analyses as ``regular analyses'' after regular languages.}
%
CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics---especially if they feature non-local control transfer that breaks the pushdown model.
%%

@@

%%
The $\Addr$ space is what controls the precision of the model.
%
For a concrete semantics, we require the allocation meta-function $\alloc :
\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any strategy is sound~\citep{dvanhorn:Might2009Posteriori}.
%
If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of continuation frames is finite.
%
A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet''---thus the pushdown system interpretation is apt.
%
The $\alloc$ meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
%
All recursion can be expressed by indirecting through addresses into a store.
%
We borrow this technique on top of our own.

@@

%%
The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
%
A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
%
In this case, the context includes the function, environment, and store.
%
All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
%%

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified thusly:

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

\begin{figure}
      \begin{tabular}{rcccl}
        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
        & $\Rightarrow$ &
        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
        & $\Rightarrow$ &
        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
      \end{tabular}
      \begin{center}
      \begin{tabular}{|l|c|c|c|}
        \hline
        \# & $\mktab$ & $\mmemo$ \\
        \hline
        0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
        1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
        \hline
        \end{tabular}
      \begin{tabular}{|l|c|}
        \hline
        \# & $\mctx$ \\
        \hline
        0  & $(({\tt x}, \iota), \mstore_1)$ \\
        1  & $(({\tt x}, \iota), \mstore_3)$ \\
        \hline
      \end{tabular}
      \begin{tabular}{|l|c|c|}
        \hline
        \# & $\mstore$ & $\mkont$ \\
        \hline
        0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
        1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
        2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
        3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
        4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
        \hline
      \end{tabular}
      \end{center}
      \caption{Example analysis evaluation}
      \label{fig:ex-eval}
\end{figure}
% %% FIXME stores in memo-table
% \begin{enumerate}
% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
% \end{enumerate}

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{center}
  \,
  \begin{minipage}{0.32\linewidth}
    $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
  \end{minipage}
  \begin{minipage}{0.36\linewidth}
    $\touchesf(\kar{\mexpr,\menv}) = \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})$
  \end{minipage}
  \begin{minipage}{0.3\linewidth}
    $\touchesf(\kfn{\mval}) = \touches(\mval)$
  \end{minipage}
\end{center}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

@@

A tabular semantics for the CES$\xi$K machine follows the previous section, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
\begin{align*}
  \tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
  \tpl{\mpoint,
    \mstore',
    \mframe',
    \krt{\mpoint, \mstore'},
    \mktab',
    \mmemo} \\
  \text{where }
  & \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
  & (\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval) \\
  & \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{(\mkont,\mframe)}
\\[2pt]
 \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
 \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
  \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
\end{align*}

%%
%% CFA2 also includes what they called ``transitive summaries'' to deal with tail calls.
%% %
%% We create an $\rt$ continuation at every function call.
%% %
%% This view contends with tail calls, but we can identify tail calls easily---any call with an $\rt$ as its continuation is a tail call.
%% %
%% Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important since there is no unbounded recursion.
%% %
%% The repeated popping of $\rt$ inserted by what otherwise were tail-calls is synonymous with CFA2's transitive summaries.

@@

Second, the $\domarkset$ metafunction builds a map of keys to lists, and builds a private store for the list constructions.
%
For this, we extend $\alloc$'s specification to allow more arguments---the private store, and a set of keys that need cons cells allocated for them---and produce a map from keys to pairs of addresses.
%
%% FIXME: need to crawl through qualified continuations for all extensions to build a list for each found key.
%% Has another implicit equality check
\newcommand{\markaux}{\mathit{marks}}
\begin{align*}
  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &= \markaux(\mkont,\emptyset) \\
  \markaux(\kmt^\bot, G) &= (\bot, \bot) \\
  \markaux(\krt{\mctx}^\mmarks, G) &= (\bot, \bot) \text{ if } \mctx \in G \\ %% FIXME: unsound
  \markaux(\krt{\mctx}^\mmarks, G) &= \bigsqcup\limits_{\mkont\in \mktab(\mctx)}{\markaux(\mkont,G \cup \set{\mctx})} \text{ otherwise} \\
  \markaux(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, G) &= (\mmarks''[\mval_k \mapsto \set{\cons{a}{d}} : \mval_k \in \dom(\mmarks), (a, d) = A(\mval_k)], \\
    &\phantom{= (}
      \mstore\sqcup[a \mapsto \mmarks(\mval_k) : \mval_k \in \dom(\mmarks), (a,\_) = A(\mval_k)] \\
    &\phantom{= (\mstore}\sqcup[d \mapsto \mmarks''(\mval_k) : \mval_k \in \dom(\mmarks), (\_,d) = A(\mval_k)]
   \\ \text{where } (\mmarks'', \mstore) &= \markaux(\mkont, G)
   \\ A &= \alloc(\mstate, \mstore, \dom(\mmarks))
\end{align*}
}

@@

At returns, the memo table must add the result to all represented contexts:
\begin{align*}
  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \mstore', \mmktab'), \mathit{out}) &=
  [((\mexpr,\menv), \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
    \text{where } C &= \setbuild{((\mexpr,\menv), \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
\end{align*}

@@

\paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
We lose precision if we use a CPS transform to compile away {\tt shift} and {\tt reset} forms, because variables are treated less precisely than continuations.
%
%In pushdown analysis, we track function return points with entire calling contexts rather than the $k$-CFA way of allocating the continuation in the heap at some lower-precision address, like a variable.
%
%In our summarizing analysis of delimited composable control, we get added precision for {\tt reset}, {\tt shift}, and continuation invocation points.
%
Consider the following program and its CPS transform for comparison:
% \begin{lstlisting}[mathescape]
% (let* ([id ($\lambda$ (x) x)]
%        [f ($\lambda$ (y) (shift k (k (k y))))]
%        [g ($\lambda$ (z) (reset (id (f z))))])
%   (<= (g 0) (g 1)))
% \end{lstlisting}
\begin{small}
\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}

@@

\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\le$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
\end{small}
PDCFA with a monovariant allocation strategy will predict the CPS'd version returns true or false.
%
Our specialized analysis for delimited control, will predict the non-CPS'd version returns true.}

@@

%%
Pushdown models have existed in the first-order static analysis literature~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
%
The important difference when we move higher-order is that the model construction to feed these methods is an additional problem---the one we solve here.
%%
